Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence

### 2019-04-25 eDriversMS-v2.Rmd
super minor comments on MS

The St. Lawrence System is a vast and complex social-ecological system providing
a wealth of ecosystem services sustaining numerous economic sectors. Related
human pressures combine with climate driven environmental changes to produce
intricate exposure regimes of overlapping and potentially interacting
environmental drivers (*e.g.* hypoxia and fisheries) threatening ecosystems.
There is a critical need to systematically characterize the distribution,
intensity and overlap between drivers to support integrative initiatives such
as ecosystem-based management. While portals providing knowledge on individual
drivers exist, platforms collating comparable and interoperable knowledges on
multiple drivers remain conspicuously missing. This paper presents two distinct,
but related objectives. The first is to characterize the distribution and
intensity of drivers in the St. Lawrence System. The second is to launch
*eDrivers*, an open knowledge platform gathering experts committed to
structuring, standardizing and sharing knowledges on drivers in support of
science and management. We gathered data on 22 coastal, climate, fisheries and
marine traffic drivers through collaborations, existing environmental
initiatives and open data portals. We show that few areas of the St. Lawrence
are free of cumulative exposure. The Estuary, the Anticosti Gyre and coastal
areas are particularly exposed, especially in the vicinity of urban centers.
We identified 6 areas of distinct cumulative exposure regime that show that
certain drivers typically co-occur in different regions of the St. Lawrence and
that coastal areas are exposed to all driver types. Of particular concern are
two threat complexes capturing most exposure hotspots that show the convergence
of contrasting exposure regimes at the head of the Laurentian Channel. These
observations are destined to improve as *eDrivers* evolves through time to
address knowledge gaps and refine current driver layers. In an effort to share
the knowledge acquired and to ensure the lasting relevance of the description
of drivers presented in this manuscript, *eDrivers* was built on a series of
guiding principles upholding existing data management and open science
standards. Ultimately, we believe that *eDrivers* represents a much needed
solution that could radically influence broad scale research and management
practices by increasing data accessibility and interoperability and by
increasing research and decision-making efficiency.

@@

The St. Lawrence System, formed by one of the largest estuaries in the world and
a vast interior sea, is a complex social-ecological system characterized by
highly variable environmental conditions and oceanographic processes, both in
space and time [@el-sabh1990a; @white1997; @dufour2007]. It thus offers a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
Lawrence System provides a wealth of ecosystem services that have historically
and contemporarily benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017 @archambault2017]. With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @governmentofquebec2015; @rqm2018], an
intensification of the human footprint is expected in the St. Lawrence System.

@@

Apart from the northeastern Gulf, the cumulative footprint of drivers is
ubiquitous in the St. Lawrence (Figure \ref{footprint}). Cumulative exposure is
generally highest along the coast (Figure \ref{footprint}), with hotposts
located in the vicinity of coastal cities (Figure \ref{hotspot}). In general,
offshore areas are less exposed to cumulative drivers, with the Estuary and the
Anticosti Gyre being notable exceptions (Figures \ref{footprint} and
\ref{hotspot}). This is not to say that offshore areas are free of exposure,
as most of the St. Lawrence is exposed to multiple overlapping drivers
(Figures \ref{footprint} and \ref{hotspot}). For example, it is worthy to note
high cumulative footprint observed at the heads of the Anticosti and Esquiman
Channels (Figure \ref{footprint}).

>>>>

The St. Lawrence System is a vast and complex social-ecological system providing
a wealth of ecosystem services sustaining numerous economic sectors. Related
human pressures combine with climate driven environmental changes to produce
intricate exposure regimes of overlapping and potentially interacting
environmental drivers (*e.g.* hypoxia and fisheries) threatening ecosystems.
There is a critical need to systematically characterize the distribution,
intensity and overlap between drivers to support integrative initiatives such
as ecosystem-based management. While portals providing knowledge on individual
drivers exist, platforms collating comparable and interoperable knowledges
<!-- this is the first time I've ever seen the plural of knowledge... technically correct, but seems weird!--> on
multiple drivers remain conspicuously missing. This paper presents two distinct,
but related objectives. The first is to characterize the distribution and
intensity of drivers in the St. Lawrence System. The second is to launch
*eDrivers*, an open knowledge platform gathering experts committed to
structuring, standardizing and sharing knowledges on drivers in support of
science and management. We gathered data on 22 coastal, climate, fisheries and
marine traffic drivers through collaborations, existing environmental
initiatives and open data portals. We show that few areas of the St. Lawrence
are free of cumulative exposure. The Estuary, the Anticosti Gyre and coastal
areas are particularly exposed, especially in the vicinity of urban centers.
We identified 6 areas of distinct cumulative exposure regime that show that
certain drivers typically co-occur in different regions of the St. Lawrence and
that coastal areas are exposed to all driver types. Of particular concern are
two threat complexes capturing most exposure hotspots that show the convergence
of contrasting exposure regimes at the head of the Laurentian Channel. These
observations are destined to improve as *eDrivers* evolves through time to
address knowledge gaps and refine current driver layers. In an effort to share
the knowledge acquired and to ensure the lasting relevance of the description
of drivers presented in this manuscript, *eDrivers* was built on a series of
guiding principles upholding existing data management and open science
standards. Ultimately, we believe that *eDrivers* represents a much needed
solution that could radically influence broad scale research and management
practices by increasing data accessibility and interoperability and by
increasing research and decision-making efficiency.

@@

The St. Lawrence System, formed by one of the largest estuaries in the world and
a vast interior sea, is a complex social-ecological system characterized by
highly variable environmental conditions and oceanographic processes, both in
space and time [@el-sabh1990a; @white1997; @dufour2007]. It thus offers a unique
and heterogeneous array of habitats suited for the establishment of diverse and
productive ecological communities [@savenkoff2000]. As a result, the St.
Lawrence System provides a wealth of ecosystem services that have historically
and contemporarily benefited the Canadian economy. It sustains a rich fisheries
industry targeting more than 50 species, serves as the gateway to eastern
North-America by granting access to more than 40 ports and the most densely
populated Canadian region, hosts a booming tourism industry and an expanding
aquaculture production, fosters emerging activities and boasts a yet untapped
hydrocarbon potential [@beauchesne2016; @schloss2017 @archambault2017].<!-- @archambault doesn't render? --> With
major investments recently made and more forthcoming in economic and
infrastructure development and research [*e.g.* @governmentofquebec2015; @rqm2018], an
intensification of the human footprint is expected in the St. Lawrence System.

@@

Apart from the northeastern Gulf, the cumulative footprint of drivers is
ubiquitous in the St. Lawrence (Figure \ref{footprint}). Cumulative exposure is
generally highest along the coast (Figure \ref{footprint}), with hotposts
located in the vicinity of coastal cities (Figure \ref{hotspot})<!-- this figure isn't loaded at the bottom -->. In general,
offshore areas are less exposed to cumulative drivers, with the Estuary and the
Anticosti Gyre being notable exceptions (Figures \ref{footprint} and
\ref{hotspot}). This is not to say that offshore areas are free of exposure,
as most of the St. Lawrence is exposed to multiple overlapping drivers
(Figures \ref{footprint} and \ref{hotspot}). For example, it is worthy to note
high cumulative footprint observed at the heads of the Anticosti and Esquiman
Channels (Figure \ref{footprint}).
